---
title: "Vignette for imagine package"
author: "Wencheng Lau-Medrano"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{imagine}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, echo = FALSE, message = FALSE}
library(imagine)
```

# imagine

**Imaging Engine, Tools for Application of Image Filters to Data Matrices**

This package was built to take numeric data matrices and apply image-filtering algoriths.

Algoriths used by `imagine` include median-filter and 2D-convolution based algoriths implemented on Rcpp (C++) because sometimes the applicaion of this filters to large numeric matrices is very intensive and time expensive.


## Installation

For installing `imagine`, as follows:

```{r, eval=FALSE}
install.packages("imagine")
```

## Engines
`imagine` performs algorithms written in Rcpp (called 'engines'), ensuring a faster application of filters. At version 1.0.0, `imagine` include two main engines, described as follows:

### Engine 1
```{Rcpp}
#include <Rcpp.h>
#include <algorithm>
#include <math.h>

using namespace Rcpp;

// [[Rcpp::export]]
NumericMatrix engine1(NumericMatrix data, NumericMatrix kernel){
  double nrows = data.nrow();
  double ncols = data.ncol();

  double knlrows = kernel.nrow();
  double knlcols = kernel.ncol();

  NumericMatrix emptyData(nrows, ncols);

  for(double j = 0; j < ncols; j++){
    for(double i = 0; i < nrows; i++){

      double cumSum = 0;
      double k = 0;

      for(double n = 0; n < knlcols; n++){
        for(double m = 0; m < knlrows; m++){
          double a = i + m - 1;
          double b = j + n - 1;

          // Multiply the value of each cell by the corresponding value of the kernel.
          if(!std::isnan(data(a, b))){
            cumSum += data(a, b)*kernel(m, n);
            k += kernel(m, n);
          }

        }
      }

      // If all values in neighborhood were NA, returns NA for this cell
      if(k < 1){
        emptyData(i, j) = NA_REAL;
      }else{
        emptyData(i, j) = cumSum/k;
      }

    }
  }

  return emptyData;
}

```

### Engine 2
```{Rcpp}
#include <Rcpp.h>
#include <algorithm>
#include <math.h>

using namespace Rcpp;

// [[Rcpp::export]]
NumericMatrix engine2(NumericMatrix data, NumericMatrix kernel, double x){
  double nrows = data.nrow();
  double ncols = data.ncol();

  double knlrows = kernel.nrow();
  double knlcols = kernel.ncol();

  NumericMatrix emptyData(nrows, ncols);
  NumericVector miniMatrix(knlrows*knlcols);

  for(double j = 0; j < ncols; j++){
    for(double i = 0; i < nrows; i++){
      for(double n = 0; n < knlcols; n++){
        for(double m = 0; m < knlrows; m++){
          double index = m*knlrows + n;
          double a = i + m - 1;
          double b = j + n - 1;

          miniMatrix[index] = data(a, b);
        }
      }

      // Sort values 
      std::sort(miniMatrix.begin(), miniMatrix.end());

      // Get value for position indicated by 'x'
      emptyData(i, j) = miniMatrix[x];
    }
  }

  return emptyData;
}

```

## Main functions

There are 3 main version at version 1.0.0, described as follows:

```{r, cache=TRUE}
# Generate example matrix
nRows <- 100
nCols <- 200

myMatrix <- matrix(runif(nRows*nCols, 0, 100), nrow = nRows, ncol = nCols)

# Add some NA random values
index <- sample(x = seq(nRows*nCols), size = as.integer(nRows*nCols*0.2), replace = FALSE)
myMatrix[index] <- NA

# Build kernel
radius <- 3
kernel <- matrix(c(2, 1, 2,
                   1, 1, 1,
                   2, 1, 2), nrow = 3)

# Apply filters
convolutionExample <- convolution2D(dataMatrix = myMatrix, kernel = kernel)
meanFilterExample <- meanFilter2D(dataMatrix = myMatrix, radius = radius)
medianFilterExample <- medianFilter2D(dataMatrix = myMatrix, radius = radius)

```

For compare results, we will plot both data (original and filtered) using `image` function, as shows in figures 1 and 2.

**Original**

```{r, message=FALSE, fig.height=3, fig.width=3, fig.cap = "Figure 2: Original matrix", results='hide', fig.pos="h", echo=FALSE}
par(mar = rep(0, 4))
image(myMatrix, zlim = c(0, 100), col = colorRampPalette(c("red", "green"))(1e3))
```


**Filtered**

```{r, message=FALSE, fig.height=2.85, fig.width=7.5, fig.cap = "Figure 1: Filtered matrices", results='hide', fig.pos="h", echo=FALSE}
par(mar = c(0, 0.5, 0, 0.5), oma = c(0, 0, 2, 0), mfrow = c(1, 3))
image(convolutionExample, zlim = c(0, 100), col = colorRampPalette(c("red", "green"))(1e3), axes = FALSE)
mtext(text = "2D convolution", side = 3, line = 0.5)

image(meanFilterExample, zlim = c(0, 100), col = colorRampPalette(c("red", "green"))(1e3), axes = FALSE)
mtext(text = "2D mean filter", side = 3, line = 0.5)

image(medianFilterExample, zlim = c(0, 100), col = colorRampPalette(c("red", "green"))(1e3), axes = FALSE)
mtext(text = "2D median filter", side = 3, line = 0.5)
```


As we can see, `NA` in borders are not well filtered. Also, the function will try to use only the values with valid information. 

## Kernel application

In the field of image processing, one of the tools most commonly used are the **convolutions**, which consist of operations between two arrays: The array of image data and small matrices, called kernels, which weights each pixel values with its corresponding neighborhood. Different kernels produce different effects, for instance blur, shifted images (right, left, up or down), sharpening, etc.

The larger the radius, more pixels remain unanalyzed at the edges. In addition, `kernel` argument of `convolution2D` is a matrix, but unlike the `meanFilter2D` and `meanFilter2D` functions, it should not necessarily have square dimensions. So, if the user specifies a rectangular kernel, the function will use the neighborhood as the Figure 3 shows. 

```{r, message=FALSE, fig.height=2, fig.width=5.7, fig.cap = "Figure 3: Neighborhood kernel application for different kernel dimensions. Black dot indicates the position of the pixel over the filter will be applied. Arrows indicates the direction of filter application", results='hide', fig.pos="h", echo=FALSE}
par(mar = rep(0, 4), mfrow = c(1, 3), xaxs = "i", yaxs = "i")

xlim <- c(-0.15, 1.1)
ylim <- c(-0.01, 1.3)

plotKernels <- function(dim1 = 3, dim2 = 3, showArrows = FALSE){
  plot(1, 1, pch = NA, axes = FALSE, xlim = xlim, ylim = ylim, xlab = NA, ylab = NA)
  polygon(x = c(0, 1, 1, 0), y = c(0, 0, 1, 1))
  
  delay1 <- ifelse(dim1 %% 2 != 0, 0, 1/(2*dim1))
  delay2 <- ifelse(dim2 %% 2 != 0, 0, 1/(2*dim2))
  
  mtext(text = paste0(dim1, "x", dim2, " kernel"), side = 3, line = -2)
  
  for(i in seq(dim1 - 1)){
    lines(c(0, 1), 1 - rep(i/dim1, 2))
  }
  
  for(i in seq(dim2 - 1)){
    lines(rep(i/dim2, 2), c(0, 1))
  }
  
  # abline(h = seq(0, 1, length.out = dim1 + 1))
  # abline(v = seq(0, 1, length.out = dim2 + 1))
  
  if(isTRUE(showArrows)){
    arrows(x0 = -0.05, y0 = 1.05, x1 = -0.05, y1 = 0.8, length = 0.1)
    arrows(x0 = -0.05, y0 = 1.05, x1 = 0.25, y1 = 1.05, length = 0.1)
  }
  
  points(0.5 - delay2, 0.5 + delay1, pch = 16)
}

# 3x3 kernel
plotKernels(3, 3, showArrows = TRUE)

# 3x4 kernel
plotKernels(3, 4)

# 4x5 kernel
plotKernels(4, 5)
```



